---
output:
  html_document: default
  pdf_document: 
    latex_engine: xelatex #pdf not working, no error either
  word_document: default
# Title: Software framework for the processing and statistical analysis of multivariate MS-data
# Owner: Laboratory of Integrative Metabolomics (LIMET), Ghent university
# Creator: Dr. Marilyn De Graeve
# Maintainer: <limet@ugent.be>
# Script: Part IV: Report_validation_statistics
---

```{r include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

```

**Report validation statistics**

This is the automatic report for the validation report part of

```{r echo = FALSE}

Date <- as.character(Sys.Date())
#cat(paste("Experiment: ", name_project, sep = ""))
#cat(paste("Date: ", Date, sep = ""))

df_info <- data.frame(col1=c("Date", "Experiment"), col2=c(Date, name_project))

suppressMessages(library(kableExtra))
kable_input <- kable(df_info)
kable_input <- kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))
gsub("<thead>.*</thead>", "", kable_input) #no header


```

<br> 



**1. Setup**

Name_group, amount_of_samples_group and samplenames_group; see samples in input sampleMetadata file defined under Projection1.

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_1Setup.txt")
INPUT <- file.path(path_data_out, INPUT)
report_setup_groups <- read.table(INPUT, header=TRUE, sep="\t")
#report_setup_groups

library(kableExtra)
kable_input <- kable(report_setup_groups)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br> 

**2. Intrument settings**

<br> 

**3. Results**

**3.1 Number of features**

Name_group, see input sampleMetadata file under Projection1
Number of features per group (mean and SD of amount_of_feat per sample), excluding values of 0.

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_3-1Numberoffeatures.txt")
INPUT <- file.path(path_data_out, INPUT)
report_features_groups <- read.table(INPUT, header=TRUE, sep="\t")
#report_features_groups

library(kableExtra)
kable_input <- kable(report_features_groups)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br>

**3.2	TIC signal**

Plot with total ion count (TIC) signal (y-axis) per sample (x-axis) with the order of the samples according to the SampleName (alphabetical).

```{r echo = FALSE, out.width = "50%", fig.align="center"}

setwd(PATH)
library(EBImage)

name_plot <- paste(name_project, "_3-2TICsignal_per_sample.png", sep="")
name_plot <- file.path(path_data_out, name_plot)
TIC <- readImage(name_plot)
display(TIC)

```

<br>

TIC statistics are mean and SD of TIC per sample in the group. The residual SD (RSD) is defined as (sd_tic / tic_mean * 100).

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_3-2TICsignal.txt")
INPUT <- file.path(path_data_out, INPUT)
report_tic_groups <- read.table(INPUT, header=TRUE, sep="\t")
#report_tic_groups

library(kableExtra)
kable_input <- kable(report_tic_groups)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br>

**3.3 PCA plotting + outlier detection**

PCAs were calculated using (<i>n</i> - 1) components. For projection1 (all groups together) that is amount of samples - 1. For individual groups, that is amount of samples in that group - 1.
PC1 (x-axis) and PC2 (y-axis) are shown below, Hotelling T^2^ 95 % confidence ellipses (black ellipse) used to calculate outliers are depicted as well. When an outlier is present, its SampleName is labelled on the PCA score plot.

PCA all groups together (Projection1)

```{r echo = FALSE, out.width = "50%", fig.align="center"}

setwd(PATH)
library(EBImage)

name_plot <- paste(name_project, "_3-3PCA_scoreplot 1ALL.png", sep="")
name_plot <- file.path(path_data_out, name_plot)
PCA <- readImage(name_plot)
display(PCA)

```

<br>

PCA per group

```{r echo = FALSE, out.width = "50%", fig.align="center"}

setwd(PATH)
library(EBImage)

files <- list.files(path=path_data_out, pattern = "_3-3PCA_scoreplot1group")

for(file_PCAgroup in files){
  name_plot <- file_PCAgroup
  name_plot <- file.path(path_data_out, name_plot)
  PCA <- readImage(name_plot)
  display(PCA)
}

```

<br>

Sample amount and names of outliers per group, calculated using Hotelling T^2^ 95 % confidence ellipses (black ellipse).

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_3-3OutliersPCA.txt")
INPUT <- file.path(path_data_out, INPUT)
report_outliers_pcas <- read.table(INPUT, header=TRUE, sep="\t")
#report_outliers_pcas

library(kableExtra)
kable_input <- kable(report_outliers_pcas)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br>

**3.4 Feature repeatability**

To evaluate the repeatability of the features, coefficient of variance (CV) values are calculated for all samples (representing biological replicates) within a group. CV values are calculated on a subset of features (features which are present in 80 % of all samples in a group, including values of zero). Because of this restriction, CV are only calculated for groups containing more than 5 samples.

<br>

**RAW preprocessed data**

CV values were calculated with raw pre-processed data (prior to any transformation / normalisation / scaling steps). CV threshold was set to 30 %. Amount and percentage of features below this threshold are listed below.

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_3-4Featurerepeatability_raw.txt")
INPUT <- file.path(path_data_out, INPUT)
report_cv_raws <- read.table(INPUT, header=TRUE, sep="\t")
#report_cv_raws

library(kableExtra)
kable_input <- kable(report_cv_raws)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br>

Per group, a cumulative histogram plot an table are shown with the amount_of_features (1st y-axis, frequency) and percentage_of_features (2nd y-axis, cumulative %) per interval (x-axis, bin). The frequency is calculated per interval (e.g., 20-30). The cumulative % is calculated from start until bin number (e.g. 0-20). For visualisation purposes, the histogram was cut-off at bin 90 (interval 90-100).

```{r echo = FALSE, out.width = "50%", fig.align="center", results='asis'}

setwd(PATH)
library(EBImage)
library(kableExtra)

files <- list.files(path=path_data_out, pattern = "_3-4Featurerepeatability_raw1group")

for(file_histgroup in files){
  name_plot <- file_histgroup
  if(substr(name_plot,(nchar(name_plot)+1)-3,nchar(name_plot)) == "png"){
    #plot histogram
    name_plot <- file.path(path_data_out, name_plot)
    PCA <- readImage(name_plot)
    display(PCA)
  }
  if(substr(name_plot,(nchar(name_plot)+1)-3,nchar(name_plot)) == "txt"){
    #table histogram
    INPUT <- file.path(path_data_out, name_plot)
    hist_table_group <- read.table(INPUT, header=TRUE, sep="\t")
    kable_input <- kable(hist_table_group)
    kable_input <- kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))
    print(kable_input)
  }
}

```

<br>

**TIC-normalized preprocessed data**

CV values were calculated with TIC-normalised data (TIC-normalisation was performed on raw pre-processed data). CV threshold was set to 30 %. Amount and percentage of features below this threshold are listed below.

```{r echo = FALSE}

setwd(PATH)

INPUT <- paste0(name_project, "_3-4Featurerepeatability_tic.txt")
INPUT <- file.path(path_data_out, INPUT)
report_cv_tics <- read.table(INPUT, header=TRUE, sep="\t")
#report_cv_tics

library(kableExtra)
kable_input <- kable(report_cv_tics)
kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))

```

<br>

Per group, a cumulative histogram plot an table are shown with the amount_of_features (1st y-axis, frequency) and percentage_of_features (2nd y-axis, cumulative %) per interval (x-axis, bin). The frequency is calculated per interval (e.g., 20-30). The cumulative % is calculated from start until bin number (e.g. 0-20). For visualisation purposes, the histogram was cut-off at bin 90 (interval 90-100).

```{r echo = FALSE, out.width = "50%", fig.align="center", results='asis'}

setwd(PATH)
library(EBImage)
library(kableExtra)

files <- list.files(path=path_data_out, pattern = "_3-4Featurerepeatability_tic1group")

for(file_histgroup in files){
  name_plot <- file_histgroup
  if(substr(name_plot,(nchar(name_plot)+1)-3,nchar(name_plot)) == "png"){
    #plot histogram
    name_plot <- file.path(path_data_out, name_plot)
    PCA <- readImage(name_plot)
    display(PCA)
  }
  if(substr(name_plot,(nchar(name_plot)+1)-3,nchar(name_plot)) == "txt"){
    #table histogram
    INPUT <- file.path(path_data_out, name_plot)
    hist_table_group <- read.table(INPUT, header=TRUE, sep="\t")
    kable_input <- kable(hist_table_group)
    kable_input <- kable_styling(kable_input, latex_options= c("scale_down", "hold_position"))
    print(kable_input)
  }
}

```

<br>

**References**

<u> TIC: </u> 

Deininger S-O, Cornett DS, Paape R, et al. Normalization in MALDI-TOF imaging datasets of proteins: practical considerations. Anal Bioanal Chem. 2011;401(1):167-181. doi:10.1007/s00216-011-4929-z

<u> PCA: </u> 

Abdi H. Partial least squares regression and projection on latent structure regression. Wiley Interdiscip Rev Comput Stat. 2010;2(1):97-106. doi:10.1002/wics.051

<u> Hotelling T^2^: </u> 

Wikström C, Albano C, Eriksson L, et al. Multivariate process and quality monitoring applied to an electrolysis process: Part I. Process supervision with multivariate control charts. Chemom Intell Lab Syst. 1998;42(1):221-231. doi:https://doi.org/10.1016/S0169-7439(98)00014-8

<u> CV: </u> 

Everitt BS. The Cambridge Dictionary of Statistics. Cambridge, U.K.; 1998.

<br>

<br>

End of report

<br>
